---
title: "Svx D90 CFUs"
author:
- name: Pablo Maldonado
  affiliation: Colorado State University
date: '2023-08-17'
output:
  html_document: default
  word_document: default
subtitle: Innate Immunity Pilot
---

```{r global options, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

## Data description

The data are collected in a spreadsheet with multiple sheets. The first sheet
(named "[x]") is used to record some metadata for the experiment, while the 
following sheets are used to record CFUs counts from the plates used for samples
from each organ, with one sheet per organ. For example, if you plated data
from both the lung and spleen, there would be three sheets in the file: one 
with the metadata, one with the plate counts for the lung, and one with the
plate counts for the spleen. 

The metadata sheet is used to record information about the overall process of
plating the data. Values from this sheet will be used in calculating the bacterial
load in the original sample based on the CFU counts. This spreadsheet includes
the following columns: 

- `organ`: Include one row for each organ that was plated in the experiment. 
You should name the organ all in lowercase (e.g., "lung", "spleen"). You 
should use the same name to also name the sheet that records data for that organ
for example, if you have rows in the metadata sheet for "lung" and "spleen", 
then you should have two other sheets in the file, one sheet named "lung" and 
one named "spleen", which you'll use to store the plate counts for each of those
organs.
- `prop_resuspended`: In this column, give the proportion of that organ that 
was plated. For example, if you plated half the lung, then in the "lung" row
of this spread sheet, you should put 0.5 in the `prop_resuspended` column. 
- `total_resuspended_uL`: This column contains an original volume of tissue homogenate. 
For example, raw lung tissue is homogenized in 500 uL of PBS in a tube containing metal beads. 
- `og_aliquot_uL`: 100 uL of th total_resuspended slurry would be considered an original 
aliquot and is used to peform serial dilutions.
- `dilution_factor`: Amount of the original stock solution that is present in the total solution, 
after dilution(s)
- `plated_uL`: Amount of suspension + diluent plated on section of solid agar 

## Read in data

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(ggpubr)
```
#Replace w/ path to CFU sheet

```{r}
path <- c("/Volumes/rstor-henao_lab/Pablo/CFU_Pipeline/IMPAcTB_Lung_Option1/cfu_revised_MHT.xlsx")
```


```{r}
sheet_names <- excel_sheets(path)
```


```{r}
merged_data <- list()

for(i in 1:length(sheet_names)){
  data <- read_excel(path, sheet = sheet_names[i]) %>% 
    mutate(organ = paste0(sheet_names[i]))
merged_data[[i]] <- data
}
  
all_data <- bind_rows(merged_data, .id = "column_label") %>% 
    dplyr::select(-column_label)
  
all_data
```

```{r}
# #Replace w/ path to CFU sheet
# all_data <- read_excel("/Users/PabloM/Downloads/cfu_revised_MHT.xlsx")
# 
# all_data
```




## Exploratory analysis and quality checks

### Exploratory analysis

**Dimensions of input data:**

Based on the input data, data were collected for the following organ or 
organs: 

The following number of mice were included for each: 

The following number of replicates were recorded at each count date for 
each experimental group: 

The following number of dilutions and dilution level were recorded for 
each organ: 

**People who plated and collected the data. Date or dates of counting:**

Based on the input data, the plates included in these data were counted by 
the following person or persons: 
Based on the input data, the plates included in these data were counted on 
the following date or dates: 

```{r}
all_data$Groups = factor(all_data$Groups, levels = c("Saline", "BCG", "ID93", "BCG+ID93"))

all_data <- all_data %>%
  mutate(Groups = recode(Groups,
                         "ID93" = "ID93+GLA-SE",
                         "BCG+ID93" = "BCG-ID93+GLA-SE"))
all_data

```



## Calculate CFUs from best dilution/Estimate bacterial load for each sample based on good dilution
```{r}

tidy_cfu_meta_joined <- all_data


tidy_cfu_meta_joined

```


```{r}
# Perform two-way ANOVA
anova_result <- aov(log10(CFU) ~ Groups, data = tidy_cfu_meta_joined %>% filter(Time_point == "90_DPI"))

# Summary of the two-way ANOVA
summary(anova_result)

# Perform Tukey's HSD test
tukey_result <- TukeyHSD(anova_result)

# Print Tukey's HSD test results
print(tukey_result)


# Load necessary libraries
library(ggplot2)

# Interaction plot
interaction.plot <- ggplot(tidy_cfu_meta_joined %>% filter(Time_point == "14_DPI"),
                           aes(x = Groups, y = log10(CFU), color = Sex, group = Sex)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  labs(title = "Sex Differences Between Groups 14_DPI",
       x = "Groups",
       y = "log10(CFU)") +
  theme_minimal()

print(interaction.plot)


```


```{r}
#USE FOR INDIVIDUAL DAYS
female_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Female" & Time_point == "14_DPI")

male_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Male" & Time_point == "14_DPI")

# FEMALE ANOVA
female_anova_result <- aov(log10(CFU) ~ Groups, data = female_df)

# Extract p-values from ANOVA
female_p_values <- summary(female_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
female_tukey_result <- TukeyHSD(female_anova_result)

# Extract p-values from Tukey's test
female_p_values <- female_tukey_result[[1]][, "p adj"]

# Specify the pairs of groups for significance bars
female_significant_comparisons <- female_p_values < 0.05  # Example: using a significance level of 0.05


# MALE ANOVA
male_anova_result <- aov(log10(CFU) ~ Groups, data = male_df)

# Extract p-values from ANOVA
male_p_values <- summary(male_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
male_tukey_result <- TukeyHSD(male_anova_result)

# Extract p-values from Tukey's test
male_p_values <- male_tukey_result[[1]][, "p adj"]


# Specify the pairs of groups for significance bars
male_significant_comparisons <- male_p_values < 0.05  # Example: using a significance level of 0.05




female_p_values 
female_significant_comparisons

male_p_values
male_significant_comparisons
```

```{r}
female_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Female" & Time_point == "56_DPI")

male_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Male" & Time_point == "56_DPI")

# FEMALE ANOVA
female_anova_result <- aov(log10(CFU) ~ Groups, data = female_df)

# Extract p-values from ANOVA
female_p_values <- summary(female_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
female_tukey_result <- TukeyHSD(female_anova_result)

# Extract p-values from Tukey's test
female_p_values <- female_tukey_result[[1]][, "p adj"]

# Specify the pairs of groups for significance bars
female_significant_comparisons <- female_p_values < 0.05  # Example: using a significance level of 0.05
female_significant_comparisons

# MALE ANOVA
male_anova_result <- aov(log10(CFU) ~ Groups, data = male_df)

# Extract p-values from ANOVA
male_p_values <- summary(male_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
male_tukey_result <- TukeyHSD(male_anova_result)

# Extract p-values from Tukey's test
male_p_values <- male_tukey_result[[1]][, "p adj"]


# Specify the pairs of groups for significance bars
male_significant_comparisons <- male_p_values < 0.05  # Example: using a significance level of 0.05
male_significant_comparisons

female_p_values 
male_p_values

```

```{r}
female_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Female" & Time_point == "90_DPI")

male_df <- tidy_cfu_meta_joined %>%
  filter(Sex == "Male" & Time_point == "90_DPI")

# FEMALE ANOVA
female_anova_result <- aov(log10(CFU) ~ Groups, data = female_df)

# Extract p-values from ANOVA
female_p_values <- summary(female_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
female_tukey_result <- TukeyHSD(female_anova_result)

# Extract p-values from Tukey's test
female_p_values <- female_tukey_result[[1]][, "p adj"]

# Specify the pairs of groups for significance bars
female_significant_comparisons <- female_p_values < 0.05  # Example: using a significance level of 0.05
female_significant_comparisons

# MALE ANOVA
male_anova_result <- aov(log10(CFU) ~ Groups, data = male_df)

# Extract p-values from ANOVA
male_p_values <- summary(male_anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
male_tukey_result <- TukeyHSD(male_anova_result)

# Extract p-values from Tukey's test
male_p_values <- male_tukey_result[[1]][, "p adj"]


# Specify the pairs of groups for significance bars
male_significant_comparisons <- male_p_values < 0.05  # Example: using a significance level of 0.05
male_significant_comparisons

female_p_values 
male_p_values
```



```{r}
# USE FOR INDIVIDUAL DAYS
# female_combined_df <- as.data.frame(female_p_values) %>%
#   rownames_to_column("group") %>%
#   left_join(as.data.frame(female_significant_comparisons) %>%
#               rownames_to_column("group"), by = "group")
# female_combined_df
# 
# male_combined_df <- as.data.frame(male_p_values) %>%
#   rownames_to_column("group") %>%
#   left_join(as.data.frame(male_significant_comparisons) %>%
#               rownames_to_column("group"), by = "group")
# male_combined_df
# 
# left_join(female_combined_df, male_combined_df, by = "group")
```

```{r}
# # Filter data for day 14
# df_day_14 <- tidy_cfu_meta_joined %>% 
#               filter(day == 14)
# 
# # Separate by sex for day 14
# df_day_14_male <- df_day_14 %>% 
#                   filter(sex == "male")
# 
# df_day_14_female <- df_day_14 %>% 
#                     filter(sex == "female")
# 
# # Filter data for day 56
# df_day_56 <- tidy_cfu_meta_joined %>% 
#               filter(day == 56)
# 
# # Separate by sex for day 56
# df_day_56_male <- df_day_56 %>% 
#                   filter(sex == "male")
# 
# df_day_56_female <- df_day_56 %>% 
#                     filter(sex == "female")
# 
# # Filter data for day 90
# df_day_90 <- tidy_cfu_meta_joined %>% 
#               filter(day == 90)
# 
# # Separate by sex for day 90
# df_day_90_male <- df_day_90 %>% 
#                   filter(sex == "male")
# 
# df_day_90_female <- df_day_90 %>% 
#                     filter(sex == "female")
# ```
# 
# ```{r}
# # Day+sex T.test
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_14_female, subset = group %in% c("BCG", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_14_female, subset = group %in% c("ID93+GLA-SE", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_14_female, subset = group %in% c("BCG-ID93+GLA-SE", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)
# ```
# 
# 
# ```{r}
# # Day T.test
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_56, subset = group %in% c("BCG", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_56, subset = group %in% c("ID93+GLA-SE", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)
# 
# # Perform t-test between treatment and saline groups
# t_test_result <- t.test(log10(CFUs_whole) ~ group, data = df_day_56, subset = group %in% c("BCG-ID93+GLA-SE", "Saline"))
# 
# # Extract p-value from t-test result
# p_value_t_test <- t_test_result$p.value
# 
# # Print the p-value
# print(p_value_t_test)

```



```{r fig.width=5}
ggplot(tidy_cfu_meta_joined, aes(x = Time_point, y = log10(CFU), fill = Groups, color = Groups)) +
geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", 
           cex = 1.5, position = position_dodge(width = .75)) + 
  geom_boxplot(aes(fill = Groups), alpha = 0.3) + 
  facet_wrap(~ Sex) + 
  scale_fill_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) + 
  scale_color_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) +
  theme_bw() + 
  theme(
        #axis.text.x = element_blank(), # Remove X-axis names
        axis.title.x = element_blank(),   
        axis.text.x = element_text(size = 12),
        #axis.title.x = element_text(size = 12, face = "bold"),
        #axis.text.y = element_blank(), # Remove axis names
        #axis.title.y = element_blank(), 
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10), 
        legend.key.size = unit(1, 'cm'), 
        plot.title = element_text(color="black", size=16, face="bold")) + 
  theme(strip.text = element_text(size=16, face = "bold")) +
  labs (y="Log10 CFU") +
  ggtitle("Option 1 Lung CFU") +
  ylim(4, 8)


ggplot(tidy_cfu_meta_joined, aes(x = Time_point, y = log10(CFU), fill = Groups, color = Groups)) +
geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", 
           cex = 1.5, position = position_dodge(width = .75)) + 
  geom_boxplot(aes(fill = Groups), alpha = 0.3) + 
  scale_fill_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) + 
  scale_color_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) +
  theme_bw() + 
  theme(
        #axis.text.x = element_blank(), # Remove X-axis names
        axis.title.x = element_blank(),   
        axis.text.x = element_text(size = 12),
        #axis.title.x = element_text(size = 12, face = "bold"),
        #axis.text.y = element_blank(), # Remove axis names
        #axis.title.y = element_blank(), 
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10), 
        legend.key.size = unit(1, 'cm'), 
        plot.title = element_text(color="black", size=16, face="bold")) + 
  theme(strip.text = element_text(size=16, face = "bold")) +
  labs (y="Log10 CFU") +
  ggtitle("Option 1 Lung CFU") +
  ylim(4, 8)

```

```{r}
ggplot(lungoption1tidy_cfu_meta_joined, aes(x = day, y = log10(CFUs_whole), fill = group, color = group)) +
geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", 
           cex = 1.5, position = position_dodge(width = .75)) + 
  geom_boxplot(aes(fill = group), alpha = 0.3) + 
  facet_wrap(~ sex) + 
  scale_fill_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) + 
  scale_color_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) +
  theme_bw() + 
  theme(
        #axis.text.x = element_blank(), # Remove X-axis names
        axis.title.x = element_blank(),   
        axis.text.x = element_text(size = 12),
        #axis.title.x = element_text(size = 12, face = "bold"),
        #axis.text.y = element_blank(), # Remove axis names
        #axis.title.y = element_blank(), 
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10), 
        legend.key.size = unit(1, 'cm'), 
        plot.title = element_text(color="black", size=16, face="bold")) + 
  theme(strip.text = element_text(size=16, face = "bold")) +
  labs (y="Log10 CFU") +
  ggtitle("Option 1 Lung CFU") +
  ylim(4.5, 6.5)


ggplot(lungoption1tidy_cfu_meta_joined, aes(x = day, y = log10(CFUs_whole), fill = group, color = group)) +
geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", 
           cex = 1.5, position = position_dodge(width = .75)) + 
  geom_boxplot(aes(fill = group), alpha = 0.3) + 
  scale_fill_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) + 
  scale_color_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) +
  theme_bw() + 
  theme(
        #axis.text.x = element_blank(), # Remove X-axis names
        axis.title.x = element_blank(),   
        axis.text.x = element_text(size = 12),
        #axis.title.x = element_text(size = 12, face = "bold"),
        #axis.text.y = element_blank(), # Remove axis names
        #axis.title.y = element_blank(), 
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 10), 
        legend.key.size = unit(1, 'cm'), 
        plot.title = element_text(color="black", size=16, face="bold")) + 
  theme(strip.text = element_text(size=16, face = "bold")) +
  labs (y="Log10 CFU") +
  ggtitle("Option 1 Lung CFU") +
  ylim(4.5, 6.5)


```

