---
title: "CFU Coding pipeline"
author:
- name: Pablo Maldonado
  affiliation: Colorado State University
date: '2024-2-28'
output:
  html_document: default
  word_document: default
subtitle: CFU plotting 
---

```{r global options, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

## Data description

The data are collected in a spreadsheet with multiple sheets. The first sheet
(named "[x]") is used to record some metadata for the experiment, while the 
following sheets are used to record CFUs counts from the plates used for samples
from each organ, with one sheet per organ. For example, if you plated data
from both the lung and spleen, there would be three sheets in the file: one 
with the metadata, one with the plate counts for the lung, and one with the
plate counts for the spleen. 

The metadata sheet is used to record information about the overall process of
plating the data. Values from this sheet will be used in calculating the bacterial
load in the original sample based on the CFU counts. This spreadsheet includes
the following columns: 

- `organ`: Include one row for each organ that was plated in the experiment. 
You should name the organ all in lowercase (e.g., "lung", "spleen"). You 
should use the same name to also name the sheet that records data for that organ
for example, if you have rows in the metadata sheet for "lung" and "spleen", 
then you should have two other sheets in the file, one sheet named "lung" and 
one named "spleen", which you'll use to store the plate counts for each of those
organs.
- `prop_resuspended`: In this column, give the proportion of that organ that 
was plated. For example, if you plated half the lung, then in the "lung" row
of this spread sheet, you should put 0.5 in the `prop_resuspended` column. 
- `total_resuspended_uL`: This column contains an original volume of tissue homogenate. 
For example, raw lung tissue is homogenized in 500 uL of PBS in a tube containing metal beads. 
- `og_aliquot_uL`: 100 uL of th total_resuspended slurry would be considered an original 
aliquot and is used to peform serial dilutions.
- `dilution_factor`: Amount of the original stock solution that is present in the total solution, 
after dilution(s)
- `plated_uL`: Amount of suspension + diluent plated on section of solid agar 

## Read in data

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(ggpubr)
library(viridis)
library(vcd)
library(ggmosaic)
library(kableExtra)
```

```{r}
#Replace w/ path to CFU sheet
path <- c("R:\\RSTOR-HenaO_lab\\Pablo\\CFU_Pipeline\\SVX_Boost2X_Lung_Day30_cfu_numbers.xlsx")

sheet_names <- excel_sheets(path)
sheet_names <- sheet_names[!sheet_names %in% c("metadata")]

merged_data <- list()

for(i in 1:length(sheet_names)){
  data <- read_excel(path, sheet = sheet_names[i]) %>% 
    mutate(organ = paste0(sheet_names[i]))
  data <- data %>% 
    #mutate(missing_col = NA) %>% 
    mutate_if(is.double, as.numeric) %>% 
    mutate_if(is.numeric, as.character) %>% 
    pivot_longer(starts_with("dil_"), names_to = "dilution",
                 values_to = "CFUs") %>% 
    mutate(dilution = str_extract(dilution, "[0-9]+"),
           dilution = as.numeric(dilution))

  merged_data[[i]] <- data
}
  
all_data <- bind_rows(merged_data, .id = "column_label") %>% 
    dplyr::select(-column_label)
  
head(merged_data)
head(all_data)
```

## Exploratory analysis and quality checks

### Exploratory analysis

**Dimensions of input data:**

Based on the input data, data were collected for the following organ or 
organs: 

The following number of mice were included for each: 

The following number of replicates were recorded at each count date for 
each experimental group: 

The following number of dilutions and dilution level were recorded for 
each organ: 

**People who plated and collected the data. Date or dates of counting:**

Based on the input data, the plates included in these data were counted by 
the following person or persons: 
Based on the input data, the plates included in these data were counted on 
the following date or dates: 

```{r}
all_data %>%
  dplyr::select(organ, who_plated, who_counted, count_date) %>%
  distinct()

all_data <- all_data %>%
  filter(CFUs != "TNTC")

all_data$group = factor(all_data$group, levels = c("Saline", "BCG 1x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG", "BCG + SolaVAX-IM-AdVAX 2x", "BCG + SolaVAX-IM-AdVAX + SolaVAX-IN-AdVAX", "BCG + SolaVAX-IN-Laporte 2x", "BCG 2x"))

# FILTERING GROUPS
# all_data <- all_data %>%
#   dplyr::filter(group %in% c("Saline", "BCG 1x", "BCG 2x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG"))
# 
# all_data$group = factor(all_data$group, levels = c("Saline", "BCG 1x", "BCG 2x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG"))

all_data <- all_data %>%
  mutate(group = recode(group,
                         "BCG + SolaVAX-IM-CpG 2x" = "BCG + SolaVAX-IM-CpG1018 2x",
                         "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG" = "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018",
                        "BCG 2x" = "BCG + BCG",
                        "BCG 1x" = "BCG",
                        "BCG + SolaVAX-IN-Laporte 2x" = "BCG + SolaVAX-Mtb + Laporte-IN-2x"))


all_data %>%
  distinct(group)


```


**Distribution of CFUs at each dilution:**

Here's a plot that shows how many plates were too numerous to count at each 
dilution level: 

Here is a plot that shows how the CFU counts were distributed by dilution
level in the data: 

## Identify a good dilution for each sample


```{r}
#choose max cfu in each sample

# 
# tidy_cfu_data <- all_data %>%
#   mutate(dilution = str_extract(dilution, "[0-9]+"),
#          dilution = as.numeric(dilution)) %>%
#   mutate(CFUs = as.numeric(CFUs)) %>%
#   filter(CFUs >= 2, CFUs <= 200) %>%
#   group_by(group, mouse, organ, sex) %>%
#   slice_max(CFUs) %>%
#   ungroup()



tidy_cfu_data <- all_data %>%
  mutate(dilution = str_extract(dilution, "[0-9]+"),
         dilution = as.numeric(dilution)) %>%
  mutate(CFUs = as.numeric(CFUs)) %>%
  filter(CFUs >= 2, CFUs <= 120) %>%
  #filter(dilution == 5) %>%
  #filter(dilution %in% c(2, 3)) %>%
  group_by(group, mouse, organ) %>%
  slice(which.min(ifelse(CFUs >= 2 & CFUs <= 120, dilution, Inf))) %>% # which.min returns the index of the row with the minimum dilution value within the specified CFUs range
  ungroup()






```


```{r}
# Calculating CFU/ml for every qualifying replicate between 10-75 CFUs. Column binding by organ name to the metadata sheet via inner_join().
meta <- read_excel(path, sheet = "metadata")

tidy_cfu_meta_joined  <- inner_join(meta, tidy_cfu_data) %>%
  group_by(group) %>% 
  mutate(CFUs_whole = ((CFUs * dilution_factor^dilution/volume_plated_ml) * total_resuspension_mL)/.33) %>%
  dplyr::select(organ, count_date, who_plated, who_counted, group,  mouse, dilution, 
         CFUs, CFUs_whole) %>% 
  ungroup()

# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "D" & group == "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018"))
# 
# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "B" & group == "BCG 1x"))

tidy_cfu_meta_joined 
```


## Calculate CFUs from best dilution/Estimate bacterial load for each sample based on good dilution
```{r}
# Calculating CFU/ml for every qualifying replicate between 10-75 CFUs. Column binding by organ name to the metadata sheet via inner_join().
meta <- read_excel(path, sheet = "metadata")

lungD30tidy_cfu_meta_joined  <- inner_join(meta, tidy_cfu_data) %>%
  group_by(group) %>% 
  mutate(CFUs_whole = ((CFUs * dilution_factor^dilution/volume_plated_ml) * total_resuspension_mL)/.33) %>%
  dplyr::select(organ, count_date, who_plated, who_counted, group,  mouse, dilution, 
         CFUs, CFUs_whole) %>% 
  ungroup()

# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "D" & group == "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018"))
# 
# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "B" & group == "BCG 1x"))

# lungD90tidy_cfu_meta_joined 
# 
# setwd("R:/RSTOR-HenaO_lab/Pablo/CFU_Pipeline")
# getwd()
# 
# write.csv(lungD90tidy_cfu_meta_joined, "lungD90tidy_cfu_meta_joined.csv", row.names = FALSE)

# Print the dataframe
lungD30tidy_cfu_meta_joined


# tidy_cfu_meta_joined <- tidy_cfu_meta_joined %>%
#   dplyr::filter(!group %in% c("BCG + SolaVAX-IM-AdVAX 2x", "BCG + SolaVAX-IM-CpG1018 2x", "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018", "BCG + SolaVAX-IM-AdVAX + SolaVAX-IN-AdVAX"))
# 
# tidy_cfu_meta_joined$group = factor(tidy_cfu_meta_joined$group, levels = c("Saline", "BCG", "BCG + BCG", "BCG + SolaVAX-Mtb + Laporte-IN-2x"))


tidy_cfu_meta_joined %>%
  distinct(group)
```
```{r}
library(knitr)
library(broom)
glm_result <- glm(CFUs_whole ~ relevel(group, ref = "BCG"), 
                  data = tidy_cfu_meta_joined, 
                  family = quasipoisson)

glm_summary <- tidy(glm_result)

# Add significance stars based on p-value
glm_summary$Significance <- with(glm_summary, 
                                 ifelse(p.value < 0.001, "***", 
                                        ifelse(p.value < 0.01, "**", 
                                               ifelse(p.value < 0.05, "*", ""))))


glm_summary$term <- gsub("relevel\\(group, ref = \"BCG\"\\)", "", glm_summary$term)  # Remove the relevel part
glm_summary$term <- gsub("\\(Intercept\\)", "BCG", glm_summary$term)


glm_summary$term <- factor(glm_summary$term, 
                           levels = c("Saline", 
                                      "BCG", 
                                      "BCG + BCG", 
                                      "BCG + SolaVAX-IM-CpG1018 2x", 
                                      "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018", 
                                      "BCG + SolaVAX-IM-AdVAX 2x", 
                                      "BCG + SolaVAX-IM-AdVAX + SolaVAX-IN-AdVAX", 
                                      "BCG + SolaVAX-Mtb + Laporte-IN-2x"))

colnames(glm_summary) <- c("Coefficient", "Estimate", "SE", "Statistic", "p.value", "Significance")
glm_summary <- glm_summary %>%
  dplyr::select(-Statistic) %>%  
  dplyr::filter(Coefficient != "BCG")


glm_summary %>%
  kable("html", caption = " ") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, font_size = 12, color = "black") %>%  # Customize header text color to black
  column_spec(1:ncol(glm_summary), 
              extra_css = "white-space: nowrap; color: black;") 
```


```{r}

# ANOVA
anova_result <- aov(log10(CFUs_whole) ~ group, data = tidy_cfu_meta_joined)

# Extract p-values from ANOVA
p_values <- summary(anova_result)[["Pr(>F)"]]

# Perform Tukey's post hoc test for group comparisons
tukey_result <- TukeyHSD(anova_result)

# Extract p-values from Tukey's test
p_values <- tukey_result[[1]][, "p adj"]

p_values
# Specify the pairs of groups for significance bars
significant_comparisons <- p_values < 0.05  # Example: using a significance level of 0.05
significant_comparisons

# Create a data frame
comparison_df <- data.frame(
  p_value = p_values,
  significant = significant_comparisons)
comparison_df <- comparison_df %>% rownames_to_column("vaccine")

comparison_df


# Create a function to add asterisks based on p-values
add_asterisks <- function(p_value) {
  if (p_value < 0.00009) {
    return("****")
  } else if (p_value < 0.0009) {
    return("***")
  } else if (p_value < 0.009) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}

# Add asterisks to the comparison_df
comparison_df$`significance level` <- sapply(comparison_df$p_value, add_asterisks)

# Display the updated comparison_df
print(comparison_df)





# Print the resulting data frame
library(kableExtra)

kable(comparison_df, caption = "Comparison Results", align = "c")
```


```{r}
# female_combined_df <- as.data.frame(female_p_values) %>%
#   rownames_to_column("group") %>%
#   left_join(as.data.frame(female_significant_comparisons) %>%
#               rownames_to_column("group"), by = "group")
# female_combined_df
# 
# male_combined_df <- as.data.frame(male_p_values) %>%
#   rownames_to_column("group") %>%
#   left_join(as.data.frame(male_significant_comparisons) %>%
#               rownames_to_column("group"), by = "group")
# male_combined_df
# 
# left_join(female_combined_df, male_combined_df, by = "group")


# 
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG 1x"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG 2x"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG + SolaVAX-IM-CpG1018 2x"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG + SolaVAX-IM-AdVAX 2x"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG + SolaVAX-IM-AdVAX + SolaVAX-IN-AdVAX"]))
# with(tidy_cfu_meta_joined, t.test(log10(CFUs_whole)[group == "Saline"], log10(CFUs_whole)[group == "BCG + SolaVAX-IN-Laporte 2x"]))


```


```{r}
# tidy_cfu_meta_joined %>%
#   ggplot(aes(x = group, y = log10(CFUs_whole), fill = group)) +
#   geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", cex = 1.5, position = position_dodge(width = .75)) +
#   geom_boxplot(aes(fill = group), alpha = 0.3) +  scale_fill_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) + scale_color_manual(values = c("red2", "olivedrab3", "green4",  "deepskyblue")) +
#   facet_wrap(~sex) +
#   theme_bw() +
#   theme(
#     axis.text.x = element_blank(), # Remove X-axis names
#     axis.text.y = element_text(size = 24),
#     axis.title.x = element_blank(),
#     axis.title.y = element_text(size = 14, face = "bold"),
#     legend.position = "right", # Set legend position to right
#     plot.title = element_blank()) +
#   theme(strip.text = element_text(size = 14, face = "bold")) +
#   labs(
#     title = paste0("SolaVAX Day 90 Lung CFU"),
#     x = "Group",
#     y = "Log10 Lung CFU",
#     fill = "Group") +
#   ylim(4, 7)

# FILTERING GROUPS
ggplot(tidy_cfu_meta_joined, aes(x = group, y = log10(CFUs_whole), fill = group)) +#, color = group)) +
geom_point(size = 5, shape = 21, alpha = 0.8, na.rm = T, colour = "black", 
           cex = 1.5, position = position_dodge(width = .75)) + 
    geom_boxplot(aes(fill = group), alpha = 0.3) + 
    scale_color_viridis_d(name = " ", alpha = 1) +  
  scale_fill_viridis_d(name = " ", alpha = 1) + 
  theme_minimal() + 
  theme(
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, hjust = 1, size = 16), 
        axis.text.y = element_text(size = 16),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 16), 
        legend.key.size = unit(1, 'cm'), 
        legend.position = ("right"),
        plot.title = element_text(color="black", size=16, face="bold")) + 
  theme(strip.text = element_text(size=16, face = "bold")) +
        scale_y_continuous(limits = function(x) c(min(x), max(x) * 1.1)) +
  labs(y = "Total CFU Log10", x = "", fill = "Vaccine", color = "Vaccine") +
  ggtitle("SolaVAX-TB 2x Boost Day 30") +
  ylim(4, 7)
```


