---
title: "CFU Coding pipeline"
author:
- name: Pablo Maldonado
  affiliation: Colorado State University
date: '2024-2-28'
output:
  html_document: default
  word_document: default
subtitle: CFU plotting 
---

```{r global options, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

## Data description

The data are collected in a spreadsheet with multiple sheets. The first sheet
(named "[x]") is used to record some metadata for the experiment, while the 
following sheets are used to record CFUs counts from the plates used for samples
from each organ, with one sheet per organ. For example, if you plated data
from both the lung and spleen, there would be three sheets in the file: one 
with the metadata, one with the plate counts for the lung, and one with the
plate counts for the spleen. 

The metadata sheet is used to record information about the overall process of
plating the data. Values from this sheet will be used in calculating the bacterial
load in the original sample based on the CFU counts. This spreadsheet includes
the following columns: 

- `organ`: Include one row for each organ that was plated in the experiment. 
You should name the organ all in lowercase (e.g., "lung", "spleen"). You 
should use the same name to also name the sheet that records data for that organ
for example, if you have rows in the metadata sheet for "lung" and "spleen", 
then you should have two other sheets in the file, one sheet named "lung" and 
one named "spleen", which you'll use to store the plate counts for each of those
organs.
- `prop_resuspended`: In this column, give the proportion of that organ that 
was plated. For example, if you plated half the lung, then in the "lung" row
of this spread sheet, you should put 0.5 in the `prop_resuspended` column. 
- `total_resuspended_uL`: This column contains an original volume of tissue homogenate. 
For example, raw lung tissue is homogenized in 500 uL of PBS in a tube containing metal beads. 
- `og_aliquot_uL`: 100 uL of th total_resuspended slurry would be considered an original 
aliquot and is used to peform serial dilutions.
- `dilution_factor`: Amount of the original stock solution that is present in the total solution, 
after dilution(s)
- `plated_uL`: Amount of suspension + diluent plated on section of solid agar 

## Read in data

```{r}
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(ggpubr)
library(viridis)
library(vcd)
library(ggmosaic)
library(kableExtra)
library(ggbeeswarm)
library(emmeans)
library(knitr)
library(broom)
```

```{r}
#Replace w/ path to CFU sheet
# path <- c("/Volumes/rstor-henao_lab/Pablo/CFU_Pipeline/SVX_Boost2X_Lung_Day90_cfu_numbers.xlsx")

path <- "SVX_Boost2X_Lung_Day30and90_cfu_numbers.xlsx"


sheet_names <- excel_sheets(path)
sheet_names <- sheet_names[!sheet_names %in% c("metadata")]

merged_data <- list()

for(i in 1:length(sheet_names)){
  data <- read_excel(path, sheet = sheet_names[i]) %>% 
    mutate(organ = paste0(sheet_names[i]))
  data <- data %>% 
    #mutate(missing_col = NA) %>% 
    mutate_if(is.double, as.numeric) %>% 
    mutate_if(is.numeric, as.character) %>% 
    pivot_longer(starts_with("dil_"), names_to = "dilution",
                 values_to = "CFUs") %>% 
    mutate(dilution = str_extract(dilution, "[0-9]+"),
           dilution = as.numeric(dilution))

  merged_data[[i]] <- data
}
  
all_data <- bind_rows(merged_data, .id = "column_label") %>% 
    dplyr::select(-column_label)
  
head(merged_data)
all_data
```

## Exploratory analysis and quality checks

### Exploratory analysis

**Dimensions of input data:**

Based on the input data, data were collected for the following organ or 
organs: 

The following number of mice were included for each: 

The following number of replicates were recorded at each count date for 
each experimental group: 

The following number of dilutions and dilution level were recorded for 
each organ: 

**People who plated and collected the data. Date or dates of counting:**

Based on the input data, the plates included in these data were counted by 
the following person or persons: 
Based on the input data, the plates included in these data were counted on 
the following date or dates: 

```{r}
all_data %>%
  dplyr::select(organ, who_plated, who_counted, count_date) %>%
  distinct()

all_data <- all_data %>%
  filter(CFUs != "TNTC")

all_data$group = factor(all_data$group, levels = c("Saline", "BCG 1x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG", "BCG + SolaVAX-IM-AdVAX 2x", "BCG + SolaVAX-IM-AdVAX + SolaVAX-IN-AdVAX", "BCG + SolaVAX-IN-Laporte 2x", "BCG 2x"))

# FILTERING GROUPS
# all_data <- all_data %>%
#   dplyr::filter(group %in% c("Saline", "BCG 1x", "BCG 2x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG"))
# 
# all_data$group = factor(all_data$group, levels = c("Saline", "BCG 1x", "BCG 2x", "BCG + SolaVAX-IM-CpG 2x", "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG"))

all_data <- all_data %>%
  mutate(group = recode(group,
                         "BCG + SolaVAX-IM-CpG 2x" = "BCG + SolaVAX-IM-CpG1018 2x",
                         "BCG + SolaVAX-IM-CpG + SolaVAX-IN-CPG" = "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018",
                        "BCG 2x" = "BCG + BCG",
                        "BCG 1x" = "BCG",
                        "BCG + SolaVAX-IN-Laporte 2x" = "BCG + SolaVAX-Mtb + Mucosimmune-IN-2x"))


all_data %>%
  distinct(group)


```

**Distribution of CFUs at each dilution:**

Here's a plot that shows how many plates were too numerous to count at each 
dilution level: 

Here is a plot that shows how the CFU counts were distributed by dilution
level in the data: 

## Identify a good dilution for each sample


```{r}
#choose max cfu in each sample

# 
# tidy_cfu_data <- all_data %>%
#   mutate(dilution = str_extract(dilution, "[0-9]+"),
#          dilution = as.numeric(dilution)) %>%
#   mutate(CFUs = as.numeric(CFUs)) %>%
#   filter(CFUs >= 2, CFUs <= 200) %>%
#   group_by(group, mouse, organ, sex) %>%
#   slice_max(CFUs) %>%
#   ungroup()


tidy_cfu_data <- all_data %>%
  mutate(dilution = str_extract(dilution, "[0-9]+"),
         dilution = as.numeric(dilution)) %>%
  mutate(CFUs = as.numeric(CFUs)) %>%
  filter(CFUs >= 2, CFUs <= 120) %>%
  #filter(dilution == 5) %>%
  #filter(dilution %in% c(2, 3)) %>%
  group_by(day, group, mouse, organ) %>%
  slice(which.min(ifelse(CFUs >= 2 & CFUs <= 120, dilution, Inf))) %>% # which.min returns the index of the row with the minimum dilution value within the specified CFUs range
  ungroup()

# Calculating CFU/ml for every qualifying replicate between 10-75 CFUs. Column binding by organ name to the metadata sheet via inner_join().
meta <- read_excel(path, sheet = "metadata")

tidy_cfu_meta_joined  <- inner_join(meta, tidy_cfu_data) %>%
  group_by(group) %>% 
  mutate(CFUs_whole = ((CFUs * dilution_factor^dilution/volume_plated_ml) * total_resuspension_mL)/.33) %>%
  dplyr::select(organ, count_date, who_plated, who_counted, day, group, mouse, dilution, 
         CFUs, CFUs_whole) %>% 
  ungroup()

# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "D" & group == "BCG + SolaVAX-IM-CpG1018 + SolaVAX-IN-CPG1018"))
# 
# tidy_cfu_meta_joined  <- tidy_cfu_meta_joined %>%
#   filter(!(mouse == "B" & group == "BCG 1x"))

tidy_cfu_meta_joined 



```

#Multiple comparisons GLM its good 
```{r}

glm_result <- glm(CFUs_whole ~ group, data = tidy_cfu_meta_joined90, family = quasipoisson)

emmeans_result <- emmeans(glm_result, ~ group)

contrast_result <- as.data.frame(pairs(emmeans_result, reverse = TRUE))

contrast_result 


contrast_result <- as.data.frame(pairs(emmeans_result, reverse = TRUE))

# Convert contrast_result to a tibble
contrast_result_tibble <- as_tibble(contrast_result)

# View the tibble
contrast_result_tibble %>%
  filter(p.value <= 0.05)

```



# GLM Day 30

```{r}

tidy_cfu_meta_joined30 <- tidy_cfu_meta_joined %>%
  dplyr::filter(day == 30)

glm_result <- glm(CFUs_whole ~ relevel(group, ref = "BCG"), 
                  data = tidy_cfu_meta_joined30, 
                  family = quasipoisson)

glm_summary <- tidy(glm_result)

# Add significance stars based on p-value
glm_summary$Significance <- with(glm_summary, 
                                 ifelse(p.value > 0.05, "ns",
                                 ifelse(p.value < 0.001, "***", 
                                        ifelse(p.value < 0.01, "**", 
                                               ifelse(p.value < 0.05, "*", "")))))


glm_summary$term <- gsub("relevel\\(group, ref = \"BCG\"\\)", "", glm_summary$term)  # Remove the relevel part
glm_summary$term <- gsub("\\(Intercept\\)", "BCG", glm_summary$term)


glm_summary  <- glm_summary  %>%
  mutate(
    `Odds Ratio` = exp(estimate),# Exponentiate the estimate to get odds ratio
    lower_ci = exp(estimate - 1.96 * std.error), # 95% CI lower bound
    upper_ci = exp(estimate + 1.96 * std.error)  # 95% CI upper bound
  )

glm_summary <- glm_summary %>%
  mutate(
    `Percent Change in Odds` = ifelse(
      `Odds Ratio` >= 1,
      (`Odds Ratio` - 1) * 100,        # Odds ratio >= 1
      -(1 - `Odds Ratio`) * 100       # Odds ratio < 1
    ),
    `Percent Change in Odds` = ifelse(
      `Percent Change in Odds` >= 0,
      paste0("+", round(`Percent Change in Odds`, 2)), # Add "+" sign for increases
      as.character(round(`Percent Change in Odds`, 2)) # Keep "-" sign for decreases
    )
  )

glm_summary <- glm_summary %>%
  dplyr::select(term, estimate, `Odds Ratio`, `Percent Change in Odds`, Significance)


colnames(glm_summary) <- c("Group", "Estimated Effect", "Odds Ratio", "Percent Change in Odds", "Significance")
glm_summary <- glm_summary %>%
  dplyr::filter(Group != "BCG")

glm_summary[,-1] <- lapply(glm_summary[,-1], function(x) if (is.numeric(x)) signif(x, 2) else x)

# glm_summary %>%
#   kable("html", caption = " ") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
#   row_spec(0, bold = TRUE, font_size = 12, color = "black") %>%  # Customize header text color to black
#   column_spec(1:ncol(glm_summary), 
#               extra_css = "white-space: nowrap; color: black;")  # Set all text to black and prevent wrapping


library(kableExtra)
library(webshot)

glm_summary$`Estimated Effect` <- signif(glm_summary$`Estimated Effect`, digits = 2)
glm_summary30 <- glm_summary

kbl(glm_summary, caption = "Day 30 Lung CFU as Compared to Single Dose BCG", align = c("l", "c", "c", "c", "c")) %>%
  kable_classic(full_width = FALSE, lightable_options = c("striped", "hover", "condensed"))




# Convert the saved HTML to PNG
table_html <- kbl(glm_summary, 
                  caption = "Day 30 Lung CFU as Compared to Single Dose BCG", 
                  align = c("l", "c", "c", "c", "c")) %>%
  kable_classic(full_width = FALSE, lightable_options = c("striped", "hover", "condensed")) %>%
  save_kable("lungglm30.html")


table_html


webshot("lungglm30.html", "lungglm30.png", zoom = 2)

# save_kable(glmtable_html, "glmtable.html")
# library(webshot)
# 
# # Convert the HTML file to an image
# webshot("glmtable.html", "glmtable.png")

```

```{r}

tidy_cfu_meta_joined90 <- tidy_cfu_meta_joined %>%
  dplyr::filter(day == 90)

glm_result <- glm(CFUs_whole ~ relevel(group, ref = "BCG"), 
                  data = tidy_cfu_meta_joined90, 
                  family = quasipoisson)

glm_summary <- tidy(glm_result)

# Add significance stars based on p-value
glm_summary$Significance <- with(glm_summary, 
                                 ifelse(p.value > 0.05, "ns",
                                 ifelse(p.value < 0.001, "***", 
                                        ifelse(p.value < 0.01, "**", 
                                               ifelse(p.value < 0.05, "*", "")))))


glm_summary$term <- gsub("relevel\\(group, ref = \"BCG\"\\)", "", glm_summary$term)  # Remove the relevel part
glm_summary$term <- gsub("\\(Intercept\\)", "BCG", glm_summary$term)


glm_summary  <- glm_summary  %>%
  mutate(
    `Odds Ratio` = exp(estimate),# Exponentiate the estimate to get odds ratio
    lower_ci = exp(estimate - 1.96 * std.error), # 95% CI lower bound
    upper_ci = exp(estimate + 1.96 * std.error)  # 95% CI upper bound
  )

glm_summary <- glm_summary %>%
  mutate(
    `Percent Change in Odds` = ifelse(
      `Odds Ratio` >= 1,
      (`Odds Ratio` - 1) * 100,        # Odds ratio >= 1
      -(1 - `Odds Ratio`) * 100       # Odds ratio < 1
    ),
    `Percent Change in Odds` = ifelse(
      `Percent Change in Odds` >= 0,
      paste0("+", round(`Percent Change in Odds`, 2)), # Add "+" sign for increases
      as.character(round(`Percent Change in Odds`, 2)) # Keep "-" sign for decreases
    )
  )

glm_summary <- glm_summary %>%
  dplyr::select(term, estimate, `Odds Ratio`, `Percent Change in Odds`, Significance)


colnames(glm_summary) <- c("Group", "Estimated Effect", "Odds Ratio", "Percent Change in Odds", "Significance")
glm_summary <- glm_summary %>%
  dplyr::filter(Group != "BCG")

glm_summary[,-1] <- lapply(glm_summary[,-1], function(x) if (is.numeric(x)) signif(x, 2) else x)

# glm_summary %>%
#   kable("html", caption = " ") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
#   row_spec(0, bold = TRUE, font_size = 12, color = "black") %>%  # Customize header text color to black
#   column_spec(1:ncol(glm_summary), 
#               extra_css = "white-space: nowrap; color: black;")  # Set all text to black and prevent wrapping




glm_summary$`Estimated Effect` <- signif(glm_summary$`Estimated Effect`, digits = 2)

kbl(glm_summary, caption = "Day 90 Lung CFU as Compared to Single Dose BCG", align = c("l", "c", "c", "c", "c")) %>%
  kable_classic(full_width = FALSE, lightable_options = c("striped", "hover", "condensed"))

glm_summary90 <- glm_summary



table_html <- kbl(glm_summary, 
                  caption = "Day 90 Lung CFU as Compared to Single Dose BCG", 
                  align = c("l", "c", "c", "c", "c")) %>%
  kable_classic(full_width = FALSE, lightable_options = c("striped", "hover", "condensed")) %>%
  save_kable("lungglm90.html")


table_html


webshot("lungglm90.html", "lungglm90.png", zoom = 2)

# save_kable(glmtable_html, "glmtable.html")
# library(webshot)
# 
# # Convert the HTML file to an image
# webshot("glmtable.html", "glmtable.png")

```


# Merged Table
```{r}
glm_summary30 <- glm_summary30 %>%
  mutate(day = "30")

glm_summary90 <- glm_summary90 %>%
  mutate(day = "90")

mergedeglm <- bind_rows(glm_summary30, glm_summary90)
mergedeglm


reshaped_data <- mergedeglm %>%
  pivot_wider(names_from = `day`, values_from = c(`Estimated Effect`, `Odds Ratio`, `Percent Change in Odds`, Significance))
 reshaped_data <- reshaped_data[c(1, 2, 4, 6, 8, 3, 5, 7, 9)] 
colnames(reshaped_data) <- gsub("_.*", "", colnames(reshaped_data))



  
# Create the kable table
table_html <- kbl(reshaped_data) %>%
  kable_classic(full_width = FALSE, lightable_options = c("striped", "hover", "condensed")) %>%
  add_header_above(c(" " = 1, 
                     "Day 30" = 4, 
                     "Day 90" = 4)) 
```


# PLOT IT
```{r fig.width=15, fig.height=9}
# tidy_cfu_meta_joined$group <- factor(tidy_cfu_meta_joined$group,
#                            levels = c("Saline",
#                                       "BCG",
#                                       "BCG + BCG",
#                                       "BCG + SolaVAX-Mtb + Mucosimmune-IN-2x"))



tidy_cfu_meta_joined %>%
  ggplot(aes(x = group, y = log10(CFUs_whole), fill = group, color = group)) +
  geom_beeswarm(size = 3, shape = 21, alpha = 0.9, na.rm = TRUE, 
                dodge.width = 0.75, color = "black") +
  facet_wrap(~day) +
  scale_color_viridis_d(name = " ", alpha = 1) +  
  scale_fill_viridis_d(name = " ", alpha = 1) + 
  theme_minimal() + 
  theme(
        axis.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90, hjust = 1, size = 16), 
        axis.text.y = element_text(size = 16),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 16), 
        legend.key.size = unit(1, 'cm'), 
        legend.position = "bottom",  # Keep only this line for legend position
        plot.title = element_text(color="black", size=16, face="bold"),
        strip.text = element_text(size=16, face = "bold")) + 
  scale_y_continuous(limits = function(x) c(min(x), max(x) * 1.1)) + 
  labs(y = "Total CFU Log10", x = "", fill = "Vaccine", color = "Vaccine") +
  ggtitle("SolaVAX-TB 2x Boost Day 90") +
  ylim(3, 6)


#Do not delete
#df = 5

```
